Extracting The Power of Dependent Types

### 2021-09-24 paper.tex
Minor typo

Meanwhile, a deep embedding defines a data type to represent the AST
of the embedded language.  This requires us to define a type system
and evaluator for the language separately.  With a dependently typed
host language the AST and the type system can be defined as a single
data structure, which is typically referred to as a tagless
interpreter~\citep{PasalicTS02,Tagless}.  The benefit of deep embedding is that we get
full access to the syntactic structure of the embedded program.  This
approach works beautifully for embedded languages with simple type
systems. However, for embedded languages with dependent types, tagless
interpreters are technically
possible~\citep{CHAPMAN200921,10.1007/978-3-540-74464-1_7}, but are
\emph{extremely} challenging to define and use in practice.
%
In particular, the mutual dependencies between the syntax for
contexts, types, and terms and the definitional equality of the object
language require large and complex inductive-inductive or
inductive-recursive to encode
\cite{10.1007/978-3-540-74464-1_7,CHAPMAN200921,10.1145/1863495.1863497}.
%
Even if we manage to give a faithful deep embedding of the language,
actually writing programs in it is still an arduous task compared
to writing programs in the meta-language directly.

>>>>

Meanwhile, a deep embedding defines a data type to represent the AST
of the embedded language.  This requires us to define a type system
and evaluator for the language separately.  With a dependently typed
host language the AST and the type system can be defined as a single
data structure, which is typically referred to as a tagless
interpreter~\citep{PasalicTS02,Tagless}.  The benefit of deep embedding is that we get
full access to the syntactic structure of the embedded program.  This
approach works beautifully for embedded languages with simple type
systems. However, for embedded languages with dependent types, tagless
interpreters are technically
possible~\citep{CHAPMAN200921,10.1007/978-3-540-74464-1_7}, but are
\emph{extremely} challenging to define and use in practice.
%
In particular, the mutual dependencies between the syntax for
contexts, types, and terms and the definitional equality of the object
language require large and complex inductive-inductive or
inductive-recursive types to encode
\cite{10.1007/978-3-540-74464-1_7,CHAPMAN200921,10.1145/1863495.1863497}.
%
Even if we manage to give a faithful deep embedding of the language,
actually writing programs in it is still an arduous task compared
to writing programs in the meta-language directly.
